---
title: "Urchins"
author: "Eric Shaphran"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(kableExtra)
```

```{r}
urchins <- read_csv(here("data","urchins.csv")) %>%
  clean_names()
```

## **Section 1**: Size comparisons, purple and red urchins
- Exploratory data visualization

```{r}
# Faceted histogram
ggplot(data = urchins, aes(x = size)) +
  geom_histogram() +
  facet_wrap(~common_name) +
  labs(x = "Urchin diameter (cm)", y = "Number of urchins", title = "Purple urchin size vs. red urchin size")

#Boxplot
ggplot(data = urchins, aes(x = common_name, y = size)) +
  geom_boxplot() +
  labs(title = "Purple urchin size vs. red urchin size", x = "Urchin species", y = "Urchin diameter (cm)")
```

- Descriptive statistics table

```{r}
#Making the table
urchin_sizes <- urchins %>%
  group_by(common_name) %>%
  summarize(mean_diameter_cm = mean(size, na.rm = TRUE),
            median_diameter_cm = median(size, na.rm = TRUE),
            sd_cm = sd(size, na.rm = TRUE),
            sample_size = n())

#Formatting the table
urchin_sizes %>%
kable(col.names = c("Species",
                    "Mean diameter (cm)",
                      "Median diameter (cm)", 
                      "Standard deviation (cm)",
                      "Sample size (n)")) %>% 
  kable_styling(full_width = FALSE)

# Calculating the % difference between the means
(6.775489-4.469797)/4.469797
```

The sampled red urchins tends to be larger than the purple urchins. The mean red urchin diameter is 2.3cm larger, or 52% larger than the mean purple urchin diameter. There is a wider spread and therefore a larger standard deviation of red urchin sizes than purple urchin sizes.

## **Section 2**: Purple urchin counts by site and year
- Create subset for purple urchins
- Make summary table with total number of purple urchins observed by site and year
- Create exploratory graph from purple urchin summary
```{r}
purple_urchins <- urchins %>%
  filter(common_name == "Purple Urchin") %>% # Filter only purple urchins 
  mutate(date_new = mdy(date)) %>% # Convert date to class "Date" with 'mdy()'
  mutate(year = year(date_new)) %>% # Create column for year
  mutate(site_full = case_when( # Create new column called site_full containing full names of sites
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef")) 

# Make purple urchin table
tot_p_urchins <- purple_urchins %>%
  select(year, site_full) %>%
  count(year, site_full) # Use 'count()' to derive total number of purple urchins at each site for each year
tot_p_urchins # View table

# Exploratory graph of purple urchin data
ggplot(data = tot_p_urchins, aes(x = year, y = n, fill = site_full)) +
  geom_col() +
  labs(x = "Year", y = "Number of Purple Urchins", title = "Purple Urchins Along Santa Barbara Coast (SHAPHRAN)") +
  theme(legend.position = "none", plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~site_full)
```
